Phylogenetics and the human microbiome.

### 2013-09-27 sbreview.tex
Thoughts about big trees.

Very large data sets.

Some groups are not interested in accuracy per se.

Alignment.
\citep{eddy1998profile}
Profile hidden Markov models.

@@

PYNAST

@@

@@

ARB software.
Phylogenetic placement.

@@

>>>>

In general, human microbiome researchers are interested in quickly doing phylogenetic inference on large data sets, and are less interested in clade-level accuracy or measures of uncertainty.
This is not unreasonable, as for many applications such as UniFrac, the tree is used as a framework to structure the data, and there is a certain amount of flexibility in that framework that will give the same results.
Furthermore, if the underlying data it 16S we can expect some tree problems even with the best methods.

Data sets can be very large.
Size of GreenGenes tree.

The alignment culture seems to be mostly around large, established alignments.
One such alignment is the SILVA alignment \citep{pruesse2007silva,quast2013silva}, which is available as part of a web tool.

Because of the high rate of insertion and deletion
OMG the SILVA alignment is mostly gap.

Fixed alignment inferred with heuristic algorithms.

Such large data

MSA tool specializing in 16S \cite{pruesse2012sina}.

PYNAST
ARB alignment.

\citep{eddy1998profile}
Profile hidden Markov models.

@@

@@

Living tree project.

@@

ARB software.
parsimony insertion.
Phylogenetic placement.

@@

How bad of tree inference can we do and still get the same trends?
This is kind of a similar question to what is driving UniFrac?
Obviously everything can matter, but what does matter?
